---
title: "Econ HW 4"
author: "Hannah Irish & Lauren Barnum"
date: "2023-05-25"
output: 
  html_document:
      code_folding: hide
---

```{r setup, include=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(here)
```

Read in data
```{r, message=FALSE}

##read in data
damages_df <- read_csv(here("data/damages.csv")) %>%
  mutate(warming2 = warming^2)
warming <- read_csv(here("data/warming.csv"))

##name columns how I want for the warming data
colnames(warming) = c("year", "baseline", "pulse")

```

Always plot your data!
```{r}

##Plot to see what we're working with
ggplot(damages_df, aes(warming,damages))+
  geom_point() +
  labs(x="Level of Global Mean Temperature Warming (C)", y="Damages ($)", title="Damages from Degree(C) Mean Temp Warning vs. Pre-Industrial")+
  theme_minimal()
```

## **Question 1**

Fit a quadratic function
```{r}

##2nd order polynomial function
damages_qm <- lm(damages ~ 0 + warming + warming2, data=damages_df)


##extract coefficients
damage_coef <- as.numeric(damages_qm$coefficients)[2]


##test that this did what we wanted it to
degrees_test <- seq(1, 10, 0.1)
test_df <- data.frame(degrees_test) %>%
  mutate(damages = damage_coef*(degrees_test^2))

# ggplot(test_df, aes(degrees_test, damages)) +
#   geom_point() +
#   theme_minimal()

```

The formula for damages as a function of warming degrees is damages = `r format(damage_coef, big.mark=",", scientific=FALSE)`*warming^2`



```{r}

##since I made the df that used the regression values, I can use that for my line
ggplot(damages_df, aes(warming,damages))+
  geom_point() +
  geom_line(data=test_df, aes(degrees_test, damages),  color="red")+
  annotate("text", x=6, y=3000000000000000, label="Regression", color="red")+
  labs(x="Level of Global Mean Temperature Warming (C)", y="Damages ($)", title="Damages from Degree(C) Mean Temp Warning vs. Pre-Industrial")+
  theme_minimal()

```

## **Question 2**

```{r}

##make a function for my damages regression
damages_func <- function(warming){
  damages = damage_coef * warming^2 
  return(damages)
}


##calculate damages from baseline and pulse and add to data frame
warming_dam <- warming %>%
  mutate(damages_baseline = damages_func(baseline), damages_pulse = damages_func(pulse))

##calculate damage difference and difference associated with each ton increase of carbon
warming_dam <- warming_dam %>%
  mutate(difference=damages_pulse-damages_baseline, per_ton = difference/35000000000)

##plot all 4 
baseline_damages <- ggplot(warming_dam, aes(year, damages_baseline)) +
  geom_point(color="blue") +
  labs(x= "Year", y="Damages ($)", title="Baseline") +
  ylim(0, 400000000000000)+
  theme_minimal()
  

pulse_damages <- ggplot(warming_dam, aes(year, damages_pulse)) +
  geom_point(color="purple") +
  labs(x= "Year", y="Damages ($)", title="Pulse") + 
  ylim(0, 400000000000000)+
  theme_minimal()


diff_damages <- ggplot(warming_dam, aes(year, difference)) +
  geom_point(color="purple") +
  labs(x= "Year", y="Difference in Damages ($)", title="Pulse-Baseline") +
  theme_minimal()


diff_damages_co2 <- ggplot(warming_dam, aes(year, per_ton)) +
  geom_point(color="blue") +
  labs(x= "Year", y="Difference in Damages ($)", title="Pulse-Baseline per ton CO2") + 
  theme_minimal()

##arrange for .html

cowplot::plot_grid(baseline_damages,pulse_damages,diff_damages,diff_damages_co2)

```

## **Question 3**


```{r}

##make discounting function for D/(1+r)^t
discounting_func <- function(amount, t, r){
  discounted_amount = amount/(1+r)^t
}

##make arguments to use in pmap()
times <- seq(1, nrow(warming_dam), 1)
rs <- seq(0.01,0.08,0.01)
carbons <- warming_dam$per_ton

##make data frame of output of pmap()
my_matrix <- as.data.frame(purrr::pmap(list(carbons, times), discounting_func, r=rs))

##rename columns with times vector bc it's ugly
colnames(my_matrix) <- times

##switch x and y
righted <- as.data.frame(t(my_matrix))

##name columns with rates 
colnames(righted) = rs

##find sums to get social cost of carbon
scc_sums <- righted %>%
  summarize(across(c(1:ncol(righted)), sum)) %>%
  pivot_longer(everything(), names_to = "discount_rate", values_to = "scc") 

##Plot
ggplot(scc_sums, aes(discount_rate, scc)) +
  geom_point(size=3)+
  labs(x="Discount Rate", y="Cost ($)", title = "Social Cost of Carbon Under Different Discount Rates")+
  theme_minimal()
```

## **Question 4**

```{r, fig.cap = "Social Cost of Carbon under different Discount rates, with the SCC calculated by the r from the Ramsey Rule as the red triangle"}

##calculate r under ramsey rule
r_ramsey = 0.001+(2*0.01)

##make x and y switched data frame from pmap output
ramsey_df <- t(as.data.frame(purrr::pmap(list(carbons, times), discounting_func, r=r_ramsey))) 

##calculate sum (scc under Ramsey Rule)
ramsey_cost=sum(ramsey_df)

##plot first plot showing ramsey 
ggplot(scc_sums, aes(discount_rate, scc)) +
  geom_point(size=3)+
  geom_point(aes(x=r_ramsey*100, y=ramsey_cost), colour="red", shape="triangle",size=3)+
  labs(x="Discount Rate", y="Cost ($)", title = "Social Cost of Carbon Under Different Discount Rates")+
  theme_minimal()



```


## **Question 5**

```{r}


replace_1.29 <- function(year, value)
{
  output = ifelse(year<=2050, value, 1.29)
  return(output)

}

damages_5 <- warming_dam %>%
  select(1, 2) %>%
  mutate(worst_case = baseline*1.5, intervention=replace_1.29(year, baseline)) %>%
  mutate(baseline = damages_func(baseline), worst_case = damages_func(worst_case), intervention = damages_func(intervention))


damages_plotable <- damages_5 %>%
  pivot_longer(cols=c(2:4), names_to = "scenario", values_to = "warming") %>%
  mutate(damages = damages_func(warming))

ggplot(damages_plotable, aes(x=year, y=damages, color=scenario))+
  geom_point() +
  scale_color_viridis_d()+
  theme_minimal()

```



```{r}

times <- seq(1, nrow(warming_dam), 1)
r_5 <- 0.02

total_d_baseline = t(as.data.frame(purrr::pmap(list(damages_5$baseline, times), discounting_func, r=r_5)))

baseline_cost=sum(total_d_baseline)

total_d_worstcase = t(as.data.frame(purrr::pmap(list(damages_5$worst_case, times), discounting_func, r=r_5)))

worstcase_cost = sum(total_d_worstcase)

total_d_intervention = t(as.data.frame(purrr::pmap(list(damages_5$intervention, times), discounting_func, r=r_5)))

intervention_cost =sum(total_d_intervention)

optionB = intervention_cost

optionA = 0.5*(baseline_cost) + 0.5*worstcase_cost

```

Pursuing option A costs $`r format(optionA, scientific=FALSE)` and pursuing option B costs $`r format(optionB,scientific=FALSE)`  

Since Option A is $`r format(optionA-optionB, scientific=FALSE)` more costly than Option B, Option B could cost this much to still be worth picking over Option A.

If society were risk averse, then the 
